Generalized Tractability

### 2023-04-03 GeneralizedTractability_230310.tex
Fred's Home Computer finished optimality for Theorem

If \eqref{eq:tract_def} holds for some $\vp$, it clearly holds for larger, $\vp$.  We are often interested in the optimal or smallest $\vp$ for which \eqref{eq:tract_def} holds.  Define the closures of the sets of parameters for which our (strong) tractability conditions hold:
\begin{equation*}
	\cp_{\textup{trct}} : = \{\vp^* : \eqref{eq:tract_def} \text{ holds }\forall \vp \in (\vp^*,\boldsymbol{\infty})\}, \qquad
	\cp_{\textup{strct}} : = \{\vp^* : \eqref{eq:strong_tract_def} \text{ holds }\forall \vp \in (\vp^*,\boldsymbol{\infty})\}.
\end{equation*}
The  set of optimal parameters is defined as all of those parameters satisfying the (strong) tractability conditions that are not greater than or equal to others:
\begin{gather}
	\label{eq:Poptdef}
	\mathcal{P}_{\textup{opt}} : = \{\vp^* \in \cp_{\textup{trct}} :  \vp^* \notin [\vp^\dagger,\boldsymbol{\infty}) \ \forall \vp^\dagger \in  \cp_{\textup{trct}} \setminus \{\vp^*\} \}, \\
	\label{eq:Psoptdef}
	\mathcal{P}_{\textup{sopt}} : = \{\vp^* \in \cp_{\textup{strct}} :  \vp^* \notin [\vp^\dagger,\boldsymbol{\infty}) \ \forall \vp^\dagger \in  \cp_{\textup{strct}} \setminus \{\vp^*\} \}.
\end{gather}

@@

\begin{theorem}\label{thm_main_strong_tract2}
A problem is strongly $T$-tractable iff there exists $\vp \in [\vzero, \boldsymbol{\infty})$ and an integer $L_{\vp} > 0$ such that
\begin{equation} \label{eq:strong_tractiff3}
     S_{\vp}:=\sup_{d \in \naturals} \sum_{i = L_{\vp}}^\infty \frac{1}{T(\lambda_{i,d}^{-1},1,\vp)} < \infty.
\end{equation}
If \eqref{eq:strong_tractiff3} holds for some $\vp$, let  $\widetilde{\cp}_{\textup{strct}} : = \{\vp^* : \eqref{eq:strong_tractiff3} \text{ holds }\forall \vp \in (\vp^*,\boldsymbol{\infty})\}$.  Then $\cp_{\textup{strct}} = \widetilde{\cp}_{\textup{strct}}$, and the set of optimal strong tractability parameters is
\[
	\cp_{\textup{sopt}} =
	\{\vp^* \in \widetilde{\cp}_{\textup{strct}} :  \vp* \notin [\vp^\dagger,\boldsymbol{\infty}) \ \forall \vp^\dagger \in  \widetilde{\cp}_{\textup{strct}} \setminus \{\vp^*\} \}.
\]
\end{theorem}

@@

\noindent This concludes the proof of the theorem.

@@

\begin{theorem}\label{thm_main_tract2}
A problem is $T$-tractable iff there exist $\vp\ge \vzero$ and a constant $L_{\vp} > 0$ such that
\begin{equation} \label{eq:tractiff4}
     S_{\vp}:=\sup_{d \in \naturals}
     \sum_{i = \lceil L_{\vp}\, T(0,d,\vp) \rceil}^\infty \frac{1}{T(\lambda^{-1}_{i,d},d,\vp)}< \infty.
\end{equation}
If \eqref{eq:tractiff4} holds for some $\vp$, let  $\widetilde{\cp}_{\textup{trct}} : = \{\vp^* : \eqref{eq:tractiff4} \text{ holds }\forall \vp \in (\vp^*,\boldsymbol{\infty})\}$, and   $\widehat{\cp}_{\textup{trct}} : = \{\vp^* : 2\vp \in \widetilde{\cp}_{\textup{trct}}\}$.  Then $\cp_{\textup{trct}} = \widetilde{\cp}_{\textup{trct}}$, and the set of optimal strong tractability parameters is
\[
\cp_{\textup{opt}} =
\{\vp^* \in \widetilde{\cp}_{\textup{trct}} :  \vp^* \notin [\vp^\dagger,\boldsymbol{\infty}) \ \forall \vp^\dagger \in  \widetilde{\cp}_{\textup{trct}} \setminus \{\vp^*\} \}.
\]\end{theorem}

@@

Let $J(\vp,d): = \lceil L_{\vp}\, T(0,d,\vp)\rceil$.  Then this upper bound on the complexity can be further bounded above by
\begin{align*}
       \comp(\varepsilon,d)
       & = T(\varepsilon^{-1},d,\vp) \Biggl [ \sum_{i=1}^{J(\vp,d) -1} \frac{1}{T(\lambda_{i, d}^{-1},d,\vp)}
       + \underbrace{\sum_{i=J(\vp,d)}^\infty \frac{1}{T(\lambda_{i, d}^{-1},d,\vp)}}_{\le S_{\vp} \text{ by \eqref{eq:tractiff4}}} \Biggr ] \\
       & \le T(\varepsilon^{-1},d,\vp)\left[ \frac{J(\vp,d)-1}{T(\lambda_{\max}^{-1},d,\vp)} + S_{\vp}\right]
        \\
       & \le T(\varepsilon^{-1},d,\vp)\left[\frac{L_{\vp}\, T(0,d,\vp)}{T(\lambda_{\max}^{-1},1,\vp)} + S_{\vp}\right] \\
       & \le\frac{L_{\vp}}{T(\lambda_{\max}^{-1},1,\vp)} \,T(\varepsilon^{-1},d,\vp) T(0,d,\vp) + S_{\vp}\, T(\varepsilon^{-1},d,\vp)\\
       & \le\frac{L_{\vp}}{T(\lambda_{\max}^{-1},1,\vp)} \,T(\varepsilon^{-1},d,\vp) T(\varepsilon^{-1},d,\vp) + S_{\vp}\, T(\varepsilon^{-1},d,\vp)\\
       & \le \frac{L_{\vp}\,K_{\vp,2}}{T(\lambda_{\max}^{-1},1,\vp)}  T(\varepsilon^{-1},d,2\vp)  + S_{\vp}\, T(\varepsilon^{-1},d,\vp)  \qquad \text{by \eqref{eq:ptauassume}}\\
       & \le \underbrace{\left[\frac{L_{\vp}\,K_{\vp,2}}{T(\lambda_{\max}^{-1},1,\vp)} + S_{\vp}\right]}_{=:C_{2\vp}} \,T(\varepsilon^{-1},d,2\vp)
\end{align*}
It follows that we have $T$-tractability, which shows sufficiency of \eqref{eq:tractiff4}. \\

@@

Note now that, for any $d\in\naturals$,
\begin{align*}
 i(\infty,d,\vp') & = \lfloor C_{\vp'}\, T(0,d,\vp')\rfloor +1 \le  \left[C_{\vp'} + \frac{1}{T(0,d,\vp')}   \right]\, T(0,d,\vp')\\
 & \le  \underbrace{\left[C_{\vp'} + \frac{1}{T(0,d,\vp')}   \right]}_{=:L_{\vp'}}\, T(0,d,\vp')
 \le \lceil L_{\vp'} T(0,d,\vp')\rceil .
\end{align*}
For this choice of $ L_{\vp'}$ we have
\[
\sup_{d\in\naturals}}\,\, \sum_{i=\lceil L(\vp') T(0,d,\vp')\rceil}^\infty \frac{1}{T(\lambda_{i,d}^{-1},d, \vp')}
\le
\sup_{d\in\naturals} \, \, \sum_{i=i (\infty,d,\vp')+1}^\infty \frac{1}{T (\lambda_{i,d}^{-1},d,\vp')} <
\infty.
\]
This yields \eqref{eq:tractiff4} with $\vp$ replaced by $\vp'$, so we see the necessity of $\eqref{eq:tractiff4}$.

@@

\noindent \textbf{Optimality:}
\fred{Not sure how to make optimality works because for sufficiency $\vp$ turns into $2\vp$.}

>>>>

If \eqref{eq:tract_def} holds for some $\vp$, it clearly holds for larger, $\vp$.  We are often interested in the optimal or smallest $\vp$ for which \eqref{eq:tract_def} holds.  Define the closures of the sets of parameters for which our (strong) tractability conditions hold:
\begin{equation*}
	\cp_{\textup{trct}} : = \{\vp^* : \eqref{eq:tract_def} \text{ holds }\forall \vp \in (\vp^*,\boldsymbol{\infty})\}, \qquad
	\cp_{\textup{strct}} : = \{\vp^* : \eqref{eq:strong_tract_def} \text{ holds }\forall \vp \in (\vp^*,\boldsymbol{\infty})\}.
\end{equation*}
The  set of optimal parameters is defined as all of those parameters satisfying the (strong) tractability conditions that are not greater than or equal to others:
\begin{gather}
	\label{eq:Poptdef}
	\mathcal{P}_{\textup{opt}} : = \{\vp^* \in \cp_{\textup{trct}} :  \vp^* \notin [\vp,\boldsymbol{\infty}) \ \forall \vp \in  \cp_{\textup{trct}} \setminus \{\vp^*\} \}, \\
	\label{eq:Psoptdef}
	\mathcal{P}_{\textup{sopt}} : = \{\vp^* \in \cp_{\textup{strct}} :  \vp^* \notin [\vp,\boldsymbol{\infty}) \ \forall \vp \in  \cp_{\textup{strct}} \setminus \{\vp^*\} \}.
\end{gather}

@@

\begin{theorem}\label{thm_main_strong_tract2}
A problem is strongly $T$-tractable iff there exists $\vp \in [\vzero, \boldsymbol{\infty})$ and an integer $L_{\vp} > 0$ such that
\begin{equation} \label{eq:strong_tractiff3}
     S_{\vp}:=\sup_{d \in \naturals} \sum_{i = L_{\vp}}^\infty \frac{1}{T(\lambda_{i,d}^{-1},1,\vp)} < \infty.
\end{equation}
If \eqref{eq:strong_tractiff3} holds for some $\vp$, let  $\widetilde{\cp}_{\textup{strct}} : = \{\vp^* : \eqref{eq:strong_tractiff3} \text{ holds }\forall \vp \in (\vp^*,\boldsymbol{\infty})\}$.  Then $\cp_{\textup{strct}} = \widetilde{\cp}_{\textup{strct}}$, and the set of optimal strong tractability parameters is
\[
	\cp_{\textup{sopt}} =
	\{\vp^* \in \widetilde{\cp}_{\textup{strct}} :  \vp^* \notin [\vp^\dagger,\boldsymbol{\infty}) \ \forall \vp^\dagger \in  \widetilde{\cp}_{\textup{strct}} \setminus \{\vp^*\} \}.
\]
\end{theorem}

@@

\noindent This concludes the proof of  Theorem \ref{thm_main_strong_tract2}.

@@

\begin{theorem}\label{thm_main_tract2}
A problem is $T$-tractable iff there exist $\vp\ge \vzero$ and a constant $L_{\vp} > 0$ such that
\begin{equation} \label{eq:tractiff4}
     S_{\vp}:=\sup_{d \in \naturals}
     \sum_{i = \lceil L_{\vp}\, T(0,d,\vp) \rceil}^\infty \frac{1}{T(\lambda^{-1}_{i,d},d,\vp)}< \infty.
\end{equation}
If \eqref{eq:tractiff4} holds for some $\vp$, let  
\[
\widetilde{\cp}_{\textup{opt}} : = \{\vp^* : \eqref{eq:tractiff4} \text{ holds }\forall \vp \in (\vp^*,\boldsymbol{\infty}) \text{ and }  \eqref{eq:tractiff4} \text{ does not hold } \forall \vp \in [\vzero,\vp^*/2)\}.
\]
Then $\cp_{\textup{opt}} \subseteq \widetilde{\cp}_{\textup{opt}}$.
\
\end{theorem}

@@

Let $J(\vp,d): = \lceil L_{\vp}\, T(0,d,\vp)\rceil$.  Then this upper bound on the complexity can be further bounded above by
\begin{align*}
       \comp(\varepsilon,d)
       & = T(\varepsilon^{-1},d,\vp) \Biggl [ \sum_{i=1}^{J(\vp,d) -1} \frac{1}{T(\lambda_{i, d}^{-1},d,\vp)}
       + \underbrace{\sum_{i=J(\vp,d)}^\infty \frac{1}{T(\lambda_{i, d}^{-1},d,\vp)}}_{\le S_{\vp} \text{ by \eqref{eq:tractiff4}}} \Biggr ] \\
       & \le T(\varepsilon^{-1},d,\vp)\left[ \frac{J(\vp,d)-1}{T(\lambda_{\max}^{-1},d,\vp)} + S_{\vp}\right]
        \\
       & \le T(\varepsilon^{-1},d,\vp)\left[\frac{L_{\vp}\, T(0,d,\vp)}{T(\lambda_{\max}^{-1},1,\vp)} + S_{\vp}\right] \\
       & \le\frac{L_{\vp}}{T(\lambda_{\max}^{-1},1,\vp)} \,T(\varepsilon^{-1},d,\vp) T(0,d,\vp) + S_{\vp}\, T(\varepsilon^{-1},d,\vp)\\
       & \le\frac{L_{\vp}}{T(\lambda_{\max}^{-1},1,\vp)} \,T(\varepsilon^{-1},d,\vp) T(\varepsilon^{-1},d,\vp) + S_{\vp}\, T(\varepsilon^{-1},d,\vp)\\
       & \le \frac{L_{\vp}\,K_{\vp,2}}{T(\lambda_{\max}^{-1},1,\vp)}  T(\varepsilon^{-1},d,2\vp)  + S_{\vp}\, T(\varepsilon^{-1},d,\vp)  \qquad \text{by \eqref{eq:ptauassume}}\\
       & \le \underbrace{\left[\frac{L_{\vp}\,K_{\vp,2}}{T(\lambda_{\max}^{-1},1,\vp)} + S_{\vp}\right]}_{=:C_{2\vp}} \,T(\varepsilon^{-1},d,2\vp)
\end{align*}
It follows that we have $T$-tractability, which shows sufficiency of \eqref{eq:tractiff4}.

@@

Note now that, for any $d\in\naturals$,
\begin{align*}
 i(\infty,d,\vp') & = \lfloor C_{\vp'}\, T(0,d,\vp')\rfloor +1 \le  \left[C_{\vp'} + \frac{1}{T(0,d,\vp')}   \right]\, T(0,d,\vp')\\
 & \le  \underbrace{\left[C_{\vp'} + \frac{1}{T(0,1,\vp')}   \right]}_{=:L_{\vp'}}\, T(0,d,\vp')
 \le \lceil L_{\vp'} T(0,d,\vp')\rceil .
\end{align*}
For this choice of $ L_{\vp'}$ we have
\[
\sup_{d\in\naturals}}\,\, \sum_{i=\lceil L(\vp') T(0,d,\vp')\rceil}^\infty \frac{1}{T(\lambda_{i,d}^{-1},d, \vp')}
\le
\sup_{d\in\naturals} \, \, \sum_{i=i (\infty,d,\vp')+1}^\infty \frac{1}{T (\lambda_{i,d}^{-1},d,\vp')} <
\infty.
\]
This yields \eqref{eq:tractiff4} with $\vp$ replaced by $\vp'$, so we see the necessity of $\eqref{eq:tractiff4}$.

@@

\noindent \textbf{Optimality:}
Finally we need to show that $\cp_{\textup{opt}} \subseteq \widetilde{\cp}_{\textup{opt}}$.  Suppose that  $\vp^* \in \cp_{\textup{opt}}$.  We show that $\vp^*$ satisfies the two conditions for membership in $\widetilde{\cp}_{\textup{opt}}$.

By the definition of  $\cp_{\textup{opt}}$ \eqref{eq:tract_def}  holds for all $\vp \in (\vp^*,\boldsymbol{\infty})$, which implies by the argument for necessity that  \eqref{eq:tractiff4} holds for all $\vp \in (\vp^*,\boldsymbol{\infty})$.  This is the first condition for membership in $\widetilde{\cp}_{\textup{opt}}$.

Also, if  $\vp^* \in \cp_{\textup{opt}}$, then \eqref{eq:tract_def} must not hold for all $\vp \in [\vzero, \vp^*)$.  By the sufficiency argument, this implies that \eqref{eq:tractiff4} cannot hold for all $\vp \in [\vp^*/2)$.  This is the second condition for membership in $\widetilde{\cp}_{\textup{opt}}$.

Thus,  $\cp_{\textup{opt}} \subseteq \widetilde{\cp}_{\textup{opt}}$.  This concludes the proof of Theorem \ref{thm_main_tract2}.
